Weak interspecific interactions in a sagebrush steppe: evidence from observations, models, and experiments

### 2017-05-30 removals/manuscript/removal_ms.tex
Added text on ARTR response to grass removal, cleaned up files

To quantify the long-term population-level consequences of permanent competitor removals, we used an IPM to project the long-term average cover of each species. We concentrated on grass species' responses to \textit{A. tripartita} removal, since the growth regressions indicated our baseline model might underestimate competitive release in this situation (our baseline vital rate models worked well for predicting \textit{A. tripartita} responses to grass removal). Using the baseline model, with no removal treatment coeffiicients, the IPM predicts little effect of \textit{A. tripartita} removal on the cover of any grass species (compare gray and brown boxes in Fig. \ref{fig:IPMresults}). In fact, projected average cover of \textit{H. comata} and \textit{P. spicata} declines slightly following \textit{A. tripartita} removal. Using a model with removal treatment coefficients, the IPM predicts little effect of shrub removal on \textit{Poa secunda} and \textit{H. comata}, but a 2-fold increase in the cover of \textit{P. spicata} (Fig. \ref{fig:IPMresults}). Running a version of the model with maximum removal treatment effects (removal coefficients set to the 95\% confidence limit furthest from zero), resulted in the same qualitative outcome, but a larger increase in \textit{P. spicata} (Fig. \ref{fig:IPMresults-maxCI}).

@@

@@

For three of the four species, we find little or no evidence that models based on observational data underestimate 
the intensity of interspecific competition. After accounting 
for local neighborhood crowding, the removal treatment effects were never significant for \textit{A. tripartita} or \textit{H. comata}; 
our baseline model, which accounts for local neighborhood composition, adequately explained vital rate responses to removal.  
\textit{P. secunda} showed mixed responses to \textit{A. tripartita} removal: removal effects were not significant at the survival stage, 
were positive at the growth stage, and negative at the recruitment stage (where the estimated effect of \textit{A. tripartita} 
on \textit{P. secunda} is positive). At the population level, models with and without removal effects projected similar patterns of 
short (Fig. \ref{fig:IBM1step}) and long-term (Fig. \ref{fig:IPMresults}) dynamics for \textit{H. comata} and \textit{A. tripartita}. 
For \textit{P. secunda}, the model with removal effects predicted higher population growth over the short but not the long-term, perhaps 
because the short-term projections mostly reflect the positive effect of \textit{A. tripartita} removal on growth, while the long-term 
equilibrium are also sensitive to the negative effects of removal on recruitment. Although the baseline model did 
not always successfully predict removal effects on vital rates, it does appear to successfully capture the population-level 
consequences of removals on these three species.

>>>>

To quantify the long-term population-level consequences of permanent competitor removals, we used an IPM to project the long-term average cover of each species. The IPM predicts little effect of \textit{A. tripartita} removal on the cover of any grass species when we simulate the baseline model with no removal treatment coeffiicients (compare gray and brown boxes in Fig. \ref{fig:IPMresults}). In fact, projected average cover of \textit{H. comata} and \textit{P. spicata} declines slightly following \textit{A. tripartita} removal. Using a model with removal treatment coefficients, the IPM predicts little effect of shrub removal on \textit{Poa secunda} and \textit{H. comata}, but a 2-fold increase in the cover of \textit{P. spicata} (Fig. \ref{fig:IPMresults}). Running a version of the model with maximum removal treatment effects (removal coefficients set to the 95\% confidence limit furthest from zero), resulted in the same qualitative outcome, but a larger increase in \textit{P. spicata} (Fig. \ref{fig:IPMresults-maxCI}).

@@

 \new{The baseline model predicts more dramatic competitive release of \textit{A. tripartita} following grass removal: mean cover increases from 5.6\% to 16.2\%. Adding removal treatment effects dramatically alters the outcome, with \textit{A. tripartita} decreasing towards extinction. Both outcomes primarily reflect parameters in the recruitment model. The baseline model predicts an increase in cover following grass removal because of the relatively strong negative effects of \textit{Poa secunda} and \textit{H. comata} on \textit{A. tripartita} recruitment (\ref{table:recruitment}). The removal effects model reverses that outcome because the estimated coefficient for the grass removal treatment effect is very large and negative. However, all of these recruitment parameters are estimated with considerable uncertainty. For example, the effect of \textit{Poa secunda} on \textit{A. tripartita} is estimated to be -0.66, but the standard deviation of the estimate is 0.33 and the 95\% credible interval stops just short of zero. The grass removal treatment effect has a credible interval that broadly overlaps zero (\ref{table:recruitment}).  }

@@

For three of the four species, we find little or no evidence that models based on observational data underestimate 
the intensity of interspecific competition. After accounting 
for local neighborhood crowding, the removal treatment effects were never significant for \textit{A. tripartita} or \textit{H. comata}; 
our baseline model, which accounts for local neighborhood composition, adequately explained vital rate responses to removal.  
\textit{P. secunda} showed mixed responses to \textit{A. tripartita} removal: removal effects were not significant at the survival stage, 
were positive at the growth stage, and negative at the recruitment stage (where the estimated effect of \textit{A. tripartita} 
on \textit{P. secunda} is positive). At the population level, models with and without removal effects projected similar patterns of 
short (Fig. \ref{fig:IBM1step}) and long-term (Fig. \ref{fig:IPMresults}) dynamics for \textit{H. comata}. \new{ For \textit{A. tripartita}, the experimental observations and short-term population projections show little response to removal, while the long-term simulations from the baseline and removal effects models suggest strong but inconsistent responses to removal. However, these long-term simulations primarily reflect the influence of recruitment parameters that are highly uncertain. We have little confidence in the recruitment model
for \textit{A. tripartita} given the species' low (a mean of 0.63 recruits per quadrat per year, compared to means of 2.6 to 5.3 for the grasses) and episodic recruitment (only 13 \textit{A. tripartita} recruits observed since we resumed annual censuses in 2007). Given this uncertainty, we put more weight on the direct observations and short-term projections showing little effect of removal on \textit{A. tripartita} populations.  }
For \textit{P. secunda}, the model with removal effects predicted higher population growth over the short but not the long-term, perhaps 
because the short-term projections mostly reflect the positive effect of \textit{A. tripartita} removal on growth, while the long-term 
equilibrium are also sensitive to the negative effects of removal on recruitment. Although the baseline model did 
not always successfully predict removal effects on vital rates, it does appear to successfully capture the population-level 
consequences of removals on these three species.
